Model-based variable selection for regression with highly correlated variables.

### 2014-08-21 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@809 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\subsection{Sub-regressions between covariates}%\label{sectionOLS}
The classical linear regression model can be written
\begin{equation}
		\boldsymbol{Y}{|\boldsymbol{X}};\boldsymbol{\beta},\sigma_Y^2 = \boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon} \label{regressionsimple}
	\end{equation}
with $\boldsymbol{X}=(\boldsymbol{X}^1,\ldots,\boldsymbol{X}^d)$ the $n\times d$ matrix of $d$ predictor variables $\boldsymbol{X}^j$ ($j=1,\ldots,d$), $\boldsymbol{Y}$ the $n\times 1$ vector of the $n$ response variables and $\boldsymbol{\varepsilon} \sim \mathcal{N}_n(\boldsymbol{0},\sigma_Y^2\boldsymbol{I})$ the centered Gaussian noise of the regression with standard deviation $\sigma_Y >0$, $\boldsymbol{I}$ denoting the identity matrix of suitable dimension. The $d\times 1$ vector $\boldsymbol{\beta}$ gathers the coefficients of the regression\footnote{Usually a constant is included as one of the regressors. For example we can take $\boldsymbol{X}^1=(1,\ldots,1)'$. The corresponding element of $\boldsymbol{\beta}$ is then the intercept $\beta_1$.}, that can be estimated by Ordinary Least Squares (\textsc{ols}): %As shown in section \ref{sectionOLS}, 
	\begin{equation}
		\hat{\boldsymbol{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y} \label{eq:OLS}
	\end{equation}
	It is an unbiased estimate with variance matrix
	\begin{equation}
		\operatorname{Var}(\hat{\boldsymbol{\beta}})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}. \label{eq:varOLS}
	\end{equation}
	This estimate requires the inversion of $\boldsymbol{X}'\boldsymbol{X}$ which can lead to great variance estimates, so unstable estimates, if it is ill-conditioned. Ill-conditioning increases when the number $d$ of covariates grows and/or when correlations within the covariates grows (in absolute value)  \cite{hoerl1970ridge}. At the limit, when $d>n$ and/or when some correlations are maximum, $\boldsymbol{X}'\boldsymbol{X}$ becomes singular. Note that $d$ and correlations are also two non unrelated factors since, in real applications, increasing $d$ makes the risk to obtain correlated covariates higher.

@@

\paragraph{Package} \textsc{CorReg} is accessible on {\sc cran}: http://cran.r-project.org/web/packages/CorReg/index.html

>>>>

\subsection{Sub-regressions between covariates}%\label{sectionOLS}
The classical linear regression model can be written
\begin{equation}
		\boldsymbol{Y}{|\boldsymbol{X}};\boldsymbol{\beta},\sigma_Y^2 = \boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon} \label{regressionsimple}
	\end{equation}
with $\boldsymbol{X}=(\boldsymbol{X}^1,\ldots,\boldsymbol{X}^d)$ the $n\times d$ matrix of $d$ predictor variables $\boldsymbol{X}^j$ ($j=1,\ldots,d$), $\boldsymbol{Y}$ the $n\times 1$ vector of the $n$ response variables and $\boldsymbol{\varepsilon} \sim \mathcal{N}_n(\boldsymbol{0},\sigma_Y^2\boldsymbol{I})$ the centered Gaussian noise of the regression with standard deviation $\sigma_Y >0$, $\boldsymbol{I}$ denoting the identity matrix of suitable dimension. The $d\times 1$ vector $\boldsymbol{\beta}$ gathers the coefficients of the regression\footnote{Usually a constant is included as one of the regressors. For example we can take $\boldsymbol{X}^1=(1,\ldots,1)'$. The corresponding element of $\boldsymbol{\beta}$ is then the intercept $\beta_1$.}, that can be estimated by Ordinary Least Squares (\textsc{ols}): %As shown in section \ref{sectionOLS}, 
	\begin{equation}
		\hat{\boldsymbol{\beta}}=\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}\boldsymbol{X}'\boldsymbol{Y} \label{eq:OLS}
	\end{equation}
	It is an unbiased estimate with variance matrix
	\begin{equation}
		\operatorname{Var}(\hat{\boldsymbol{\beta}})=\sigma_Y^2\left(\boldsymbol{X}'\boldsymbol{X} \right) ^{-1}. \label{eq:varOLS}
	\end{equation}
	This estimate requires the inversion of $\boldsymbol{X}'\boldsymbol{X}$ which can lead to great variance estimates, so unstable estimates, if it is ill-conditioned. Ill-conditioning increases when the number $d$ of covariates grows and/or when correlations within the covariates grow (in absolute value)  \cite{hoerl1970ridge}. At the limit, when $d>n$ and/or when some correlations are maximum, $\boldsymbol{X}'\boldsymbol{X}$ becomes singular. Note that $d$ and correlations are also two non unrelated factors since, in real applications, increasing $d$ makes the risk to obtain correlated covariates higher.

@@

\paragraph{Package} {\tt CorReg} is accessible on {\sc cran}: http://cran.r-project.org/web/packages/CorReg/index.html
